---
title: "Fundamental Interactions"
---

## How This Tutorial Works

::: {.column-margin}

:::: reading

[R4DS 2.1: Coding Basics](https://r4ds.hadley.nz/workflow-basics.html#coding-basics)

::::

:::

Throughout this module, you'll move through the same four-step cycle for almost everything you learn:

1. **Do** — run a small piece of R code.
1. **See** — look carefully at what R prints back.
1. **Predict** — before you run the *next* piece of code, guess what will happen.
1. **Explain** — put into words *why* R behaved the way it did.

This cycle might feel slow at first, but it's the fastest way to build real understanding rather than just copying
commands. Let's try it right now.

## R as a Calculator

At its simplest, R is a very over-powered pocket calculator. Run the code below (click the small "Run Code" button, or
press `Ctrl`/`Cmd` + `Enter`) and look at the output.

```{webr-r}
2 + 2
```

That's the whole cycle in miniature: you *did* something (ran `2 + 2`), and you *saw* something (the number `4`).
R evaluated the expression and printed the result straight back to you.

The usual arithmetic operators all behave the way you'd expect.

```{r}
## Add 2 and 2
2 + 2

## Subtract 6 from 14
14 - 6

## Multiply 3 and 4
3 * 4

## Divide 225 by 15
225 / 15
```

::: practice

## Practice: Predict, Then Run

Before you touch the editor, **predict** what value each of the following expressions will produce. Write your guess
down (even just in your head), *then* run the code to check yourself.

```{r, eval = FALSE}
10 - 3
4 * 4
100 / 4
```

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
10 - 3
4 * 4
100 / 4
```

## {{< var tab_title.solution >}}

If your predictions matched, great! You've already got the core idea. If not, that's completely normal at this stage: notice *which* line surprised you and re-read the operator (`-`, `*`, `/`) involved. Getting comfortable naming *why* an answer is what it is (not just *that* it's correct) is exactly the skill this module is trying to build.

::::

:::

## A Slightly Bigger Task

Now let's use these same tools to solve a tiny real problem, so you can see arithmetic doing something useful rather
than just producing numbers in isolation.

::: practice

## Practice

1. Calculate the number of whole days left in the current month.
1. Calculate the number of weeks left in the month.
   - Account for partial weeks with a decimal-valued answer.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.hint >}}

Start by figuring out (in your head or on paper) how many days the current month has in total, and how many days have
already passed. Subtracting those two numbers gets you the days remaining. From there, think about which arithmetic
operator turns a number of days into a number of weeks.

## {{< var tab_title.solution >}}

At time-of-writing, it's 18 July. Since July has 31 days, we get `r 31 - 18` days remaining in the month.

```{r}
31 - 18
```

To find the number of remaining weeks, I'll just divide the answer from above by 7.

```{r}
13 / 7
```
::::

:::

Once you're confident with the ideas above, head to the [Basic Operators](/tutorials/basic_commands/basic_operators.qmd)
page to build on this foundation with all of R's arithmetic operators.
