---
title: "Understanding Arguments"
---

::: {.column-margin}

:::: reading

[R4DS 2.4: Calling Functions](https://r4ds.hadley.nz/workflow-basics.html#calling-functions)

::::

:::

## What Are Arguments?

On the last page, you already changed the inputs to `round()` and saw the output change. Those inputs have a name:
**arguments**. Arguments are simply the inputs you give to a function, written inside its parentheses.

> `function_name(arguments)`

Most R functions accept more than one argument, and each argument has its own name. You specify argument values by
writing a comma-separated list of `name = value` pairs between the parentheses of the function call.

```{r}
round(x = 3.14159, digits = 2)
```

**Do:** run the two lines of code.

```{webr-r}
round(x = 3.14159, digits = 2)
round(3.14159, 2)
```

**See:** both lines produce the same result.

**Explain:** if you supply argument values *in the order the function expects them*, you don't have to write out the
argument names. Both `round(x = 3.14159, digits = 2)` and `round(3.14159, 2)` tell R exactly the same thing — the
second version is just a shortcut.

::: practice

## Practice: Predict, Then Run

The `mean()` function accepts an argument called `trim`, which excludes a proportion of the most extreme values before
averaging. Predict what each line below will print.

```{r, eval = FALSE}
x <- c(1, 2, 3, 4, 100)
mean(x)
mean(x, trim = 0.2)
```

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
x <- c(1, 2, 3, 4, 100)
mean(x)
mean(x, trim = 0.2)
```

## {{< var tab_title.solution >}}

```{r}
x <- c(1, 2, 3, 4, 100)
mean(x)
mean(x, trim = 0.2)
```

`mean(x)` averages all five values, including the outlier `100`, which pulls the average up substantially. `mean(x,
trim = 0.2)` first removes the most extreme 20% of values on each end before averaging, giving a result that's much
closer to the "typical" value in the data.
::::

:::

## A Note on Named vs. Positional Arguments

You don't always need to write out an argument's name, but if you skip the name, R will assume you're supplying the
arguments in the order the function expects. Mixing these two styles is common and perfectly fine, as long as you keep
the ordering rules straight:

- If you name **all** your arguments, you can put them in any order.
- If you name **none** of your arguments, you must put them in the function's expected order.
- If you name **some**, the unnamed ones must still come first, in the expected order.

We'll come back to exactly how you find out a function's expected argument order and default values once we cover
[documentation](/tutorials/help/index.qmd) later in this module. For now, it's enough to recognize that arguments have
names, and that those names matter.

## Quick Check

- What is an argument?
- In `round(3.14159, 2)`, which part is the function name, and which part(s) are the arguments?
- Rewrite `round(3.14159, 2)` using explicit argument names.

```{webr-r}
```
