---
title: "Common Mistakes"
---

Everyone runs into the same handful of errors when they're getting started with functions. Recognizing them quickly
will save you a lot of frustration. Each example below is broken on purpose, try to spot the problem *before* reading
the explanation.

## Mistake 1: Forgetting a Closing Parenthesis

```{r, error = TRUE}
round(3.14159, 2
```

**Explain:** every opening parenthesis needs a matching closing parenthesis. R will often show a `+` prompt (instead
of an error) and appear to "hang" while it waits for you to finish the expression. If that happens, you're almost
always just missing a closing `)`.

## Mistake 2: Misspelling a Function Name

```{r, error = TRUE}
maen(1:5)
```

**Explain:** R is case-sensitive and doesn't guess at typos. `maen()` doesn't exist, so R tells you it "could not find
function." Double-check your spelling (and capitalization) whenever you see this error.

## Mistake 3: Misnaming an Argument

```{r, error = TRUE}
round(3.14159, digit = 2)
```

**Explain:** the `round()` function expects an argument called `digits` (plural), not `digit`. R doesn't recognize
`digit` as a valid argument name for `round()`, so it raises an error rather than guessing what you meant.

## Mistake 4: Forgetting Quotes Around Text

```{r, error = TRUE}
x <- hello
```

**Explain:** without quotation marks, R assumes `hello` is the name of an existing object, and complains that it can't
find one. To create a piece of text (a *character string*), you must wrap it in quotes: `x <- "hello"`.

::: practice

## Practice: Diagnose the Error

Each line below contains exactly one mistake from the list above (or no mistake at all). Predict what's wrong with
each one, then check the solution.

```{r, eval = FALSE}
sum(1, 2, 3
sqrt(16)
Round(2.71828, 2)
mean(x, na.rm = "yes please")
```

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
sum(1, 2, 3
```

```{webr-r}
Round(2.71828, 2)
```

## {{< var tab_title.solution >}}

- `sum(1, 2, 3` is missing a closing parenthesis.
- `sqrt(16)` has no mistake — it runs fine!
- `Round(2.71828, 2)` misspells `round()` with a capital `R`. R function names are case-sensitive, so `Round` isn't
  recognized.
- `mean(x, na.rm = "yes please")` supplies the wrong *type* of value for `na.rm`, which expects `TRUE` or `FALSE`, not
  a piece of text. We'll cover exactly how you find out what type of value an argument expects once we look at
  [documentation](/tutorials/help/index.qmd) later in this module.
::::

:::

## Quick Check

Which of the following would cause an error, and why?

- `Sum(1:5)`
- `sum(1:5`
- `sum(1:5)`
- `sum(nums = 1:5)`

Test each one in the editor to check your reasoning.

```{webr-r}
```
