---
title: "R Functions"
---

## Why Are We Programming?

Suppose you want to calculate the average of the first five positive integers, $\{1, 2, 3, 4, 5\}$. You could
directly apply the formula for the mean using nothing but the arithmetic operators you already know.

```{r}
(1 + 2 + 3 + 4 + 5) / 5
```

That works fine for five numbers. But what if you had 500 numbers? Or 5 million? Manually typing out every value and
dividing by hand simply isn't realistic. This is exactly why we program: so that a computer can do the repetitive,
error-prone parts of a calculation for us, reliably and instantly, no matter how much data we throw at it.

**Do:** run the code below.

```{webr-r}
mean(1:5)
```

**See:** it prints `3`, the same answer as our hand-calculated formula above.

**Explain:** let's break down exactly what happened in that one short line of code.

- `1:5` creates the sequence of numbers `1, 2, 3, 4, 5`.
- `mean(...)` takes that sequence and calculates its average.

So `mean(1:5)` reads almost like English: "the mean of one-through-five." Instead of writing out the entire formula
for a mean every time, we can reuse a small, pre-built piece of code, a **function**, that already knows how to do
that calculation for us.

::: practice

## Practice: Predict, Then Run

Before running the code below, predict what value each line will produce.

```{r, eval = FALSE}
1:5
mean(1:5)
sum(1:5)
```

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
1:5
mean(1:5)
sum(1:5)
```

## {{< var tab_title.solution >}}

```{r}
1:5
mean(1:5)
sum(1:5)
```

`1:5` just creates the five numbers themselves. Nothing is calculated yet. `mean(1:5)` and `sum(1:5)` both take that
same sequence of numbers as an input, but each performs a *different* operation on it: one averages the values, and
the other adds them all together. Same input, different function, different result.
::::

:::

Over the rest of this tutorial, you'll learn what a function actually *is*, how to call one, how to change what you
feed into it, and how to avoid (and fix) the most common mistakes people make when they're getting started.

## Quick Check

In your own words: why is it useful to have a function like `mean()` instead of always writing out the full averaging
formula by hand?
