---
title: "Try It"
---

Reading about functions only gets you so far. The fastest way to really understand them is to call a lot of them and
watch what happens. This page is entirely hands-on.

## Calling Your First Functions

**Do:** run each of the following function calls one at a time. For each one, pause and **see** what it prints before
moving to the next line.

```{webr-r}
sum(1:10)
```

```{webr-r}
max(c(3, 7, 2, 9, 4))
```

```{webr-r}
min(c(3, 7, 2, 9, 4))
```

```{webr-r}
length(c(3, 7, 2, 9, 4))
```

::: practice

## Practice: Change the Input

Now it's your turn. In the editor below, call `sum()`, `mean()`, and `length()` on the vector `c(10, 20, 30, 40, 50)`.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
sum(c(10, 20, 30, 40, 50))
mean(c(10, 20, 30, 40, 50))
length(c(10, 20, 30, 40, 50))
```
::::

:::

## Same Function, Different Input

One of the most important habits to build is *changing the input* to a function you already know, just to see what
happens. This is one of the best ways to build intuition for how a function behaves.

::: practice

## Practice: Predict, Then Run

Predict what each of the following will print, *before* running the code. Each line calls the same function,
`round()`, with a different input.

```{r, eval = FALSE}
round(3.14159)
round(3.14159, 2)
round(7)
round(-2.5)
```

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
round(3.14159)
round(3.14159, 2)
round(7)
round(-2.5)
```

## {{< var tab_title.solution >}}

```{r}
round(3.14159)
round(3.14159, 2)
round(7)
round(-2.5)
```

Changing the first input changes *what* gets rounded, while changing the second input (when we provide one) changes
*how precisely* it gets rounded. Notice that `round(7)` still works even though `7` is already a whole number. R
doesn't complain, it just returns `7`. We'll formally introduce these "extra" inputs, called **arguments**, on the
next page.
::::

:::

## Quick Check

Using the editor below, call `sqrt()` on three *different* numbers of your choosing. What pattern do you notice about
how the output changes as your input changes?

```{webr-r}
```
